Looking for CDI Bias: WG Comprehension

### 2021-07-31 02-preliminary-WGcomp.Rmd
new comprehension fits, multigroup model comparisons

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
library(GGally)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

@@

# mother's years of education = proxy for SES
table(d_demo$mom_ed) %>% kable()
# skewed: only 72 with Primary/Some Secondary (most have Some College +)

@@

```{r get-wordbank-items, eval=F, echo=F}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(produces = as.numeric(value == "yes" | value == "produces")) # we'll start with production
# yes = understands; no = does not understand/produce, produces = produces (and understands)

@@

d_mat <- d_long_wg %>% select(data_id, definition, produces) %>% 
  pivot_wider(id_cols = data_id, names_from = definition, values_from = produces)

save(d_mat, d_demo, en_wg, file=here("data/en_wg_comp.Rdata"))
```

@@

table(en_wg$category) %>% kable()
```

@@

@@

# ToDo: define ses_group ("high" for mom_ed > X, "low") (choose X ..high school?)

mg1 <- multipleGroup(d_mat, 1, group = sex_group) # this example should work, according to docs
mg1 <- multipleGroup(d_mat, 1, group = sex_group, 
                     invariance = c("free_var", "free_means")) 
# Error: Multiple Group model will not be identified without proper constraints (groups contain missing data patterns where item responses have been completely omitted or, alternatively, the number of categories within each group is not equal to the total number of categories)

# unless there are possibly items

@@

# constrain slopes, allow varying group means and variances -- very slow!
#mg4 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
#                     invariance =c('slopes','free_means', 'free_var'), verbose = T)

@@

anova(mod_sc1, mod_fullconstrain) # mod_fullconstrain preferred by AIC/BIC
anova(m_sc2, mod_fullconstrain) # mod_fullconstrain preferred by AIC/BIC

anova(mg2, mod_fullconstrain) # AIC prefers slope-invariant model; BIC prefers fully-constrained model

@@

anova(mg3, mg2) # AIC//BIC prefer model with varying group means
#anova(mg3, mg4)

coef(mg3)[[2]]$GroupPars # Female intercept = 0.22 (Male = 0)

@@

save(file = here("data/eng_wg_comp_sex_sexDIF.Rds"), 
     mg2, mg3,  m_sc2, mod_sc1, mod_fullconstrain) # not run: mg4
```

@@

# same item estimates with/without free_means? 
cor(mg2_df$d_f, mg3_df$d_f) # .9999
cor(mg2_df$d_m, mg3_df$d_m) # .9998

# same difficulties:
mod_sc1_df = extract_group_df(mod_sc1)

# fully-constrained is just the same as ordinary 2PL model, no? but we can use this model for DIF
mod_fullconstrain_df = extract_group_df(mod_fullconstrain)
# DIF(mod_fullconstrain, which.par = 'd')

#mod_int_df = extract_group_df(m_int)

mod_sc2_df = extract_group_df(m_sc2)

#anova(m1, mg2) # AIC prefers group model, BIC prefers 2PL
```

@@

@@

d_demo <- d_demo %>% left_join(fscores_mg2 %>% 
                                 mutate(data_id = as.numeric(data_id)))

@@

## DIF Model: SES

@@

#mg1 <- multipleGroup(d_mat, 1, group = ses_group, SE = T) 
mg2 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)

mg3 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = c('slopes', 'free_means'), verbose = T)

anova(mg2, mg3) # AIC/BIC prefer model with varying group mean

@@

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain)
```

@@

coef(mg3)[[2]]$GroupPars # low-SES group (high is ref) intercept = .24
```

@@

@@

```{r, fig.width=8, fig.height=6}
ggplot(mg3_df, 
       aes(x = d_lo, y = d_hi)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg3_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for low-SES") + 
  ylab("Item Ease for high-SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

>>>>

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
library(GGally)
library(ggpubr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

@@

@@

```{r get-wordbank-items, eval=F, echo=F}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(understands = as.numeric(value == "yes" | value == "produces" | value == "understands")) # we'll start with production
# yes = understands; no = does not understand/produce, produces = produces (and understands)

@@

d_mat <- d_long_wg %>% select(data_id, definition, understands) %>% 
  pivot_wider(id_cols = data_id, names_from = definition, values_from = understands)

d_long_wg <- d_long_wg %>% select(-value, -num_item_id)

save(d_mat, d_demo, en_wg, d_long_wg, file=here("data/en_wg_comp.Rdata"))
```

@@

table(en_wg$category) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

@@

## GLIMMER Plots

```{r fit-rasch-group-models, echo=F, eval=F}
# first fit Rasch model to each group, and then compare resulting parameters

m1pl_m <- mirt(d_mat[which(d_demo$sex=="Male"),], 1, "Rasch", SE=T, SE.type='Oakes', verbose=T)
m1pl_f <- mirt(d_mat[which(d_demo$sex=="Female"),], 1, "Rasch", SE=T, SE.type='Oakes', verbose=T)

m1pl_h <- mirt(d_mat[which(d_demo$ses_group=="high"),], 1, "Rasch", SE.type='Oakes', verbose=T)
m1pl_l <- mirt(d_mat[which(d_demo$ses_group=="low"),], 1, "Rasch", SE.type='Oakes', verbose=T)

save(file = here("data/eng_wg_rasch_group_models.Rds"),
     "m1pl_m", "m1pl_f", "m1pl_h", "m1pl_l")

res_m = residuals(m1pl_m)
#m1pl_m_ci = boot.mirt(m1pl_m)
#itemplot(m1pl_m, 1, CE=T)
```

```{r, echo=F}
load(here("data/eng_wg_rasch_group_models.Rds"))

# coef(m1pl_m) shows CI_2.5 and CI_97.5 for d...iterate over and add to dataframes?
#ci_lo = coef(m1pl_m)$`baa baa`['CI_2.5','d']
#ci_hi = coef(m1pl_m)$`baa baa`['CI_97.5','d']

par_m = as_tibble(coef(m1pl_m, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_m, simplify = TRUE)$items))
par_f = as_tibble(coef(m1pl_f, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_f, simplify = TRUE)$items))

par_h = as_tibble(coef(m1pl_h, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_h, simplify = TRUE)$items))
par_l = as_tibble(coef(m1pl_l, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_l, simplify = TRUE)$items))

rasch_pars = tibble(definition = par_m$definition,
                    d_Male = par_m$d,
                    d_Female = par_f$d,
                    d_LowSES = par_l$d,
                    d_HighSES = par_h$d) %>%
  mutate(d_sex_diff = abs(d_Female - d_Male),
         d_ses_diff = abs(d_HighSES - d_LowSES))

sex_thresh = median(rasch_pars$d_sex_diff) + sd(rasch_pars$d_sex_diff)
ses_thresh = median(rasch_pars$d_ses_diff) + sd(rasch_pars$d_ses_diff)

p1 <- ggplot(rasch_pars, 
       aes(x = d_Male, y = d_Female)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(rasch_pars, d_sex_diff > sex_thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')

p2 <- ggplot(rasch_pars, 
       aes(x = d_LowSES, y = d_HighSES)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(rasch_pars, d_ses_diff > ses_thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Low SES") + 
  ylab("Item Ease for High SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

```{r, echo=F, fig.width=10, fig.height=5}
ggarrange(p1, p2)
```

@@

# won't work if some items are all 0 (or 1?) for one sex
mg1 <- multipleGroup(d_mat, 1, group = sex_group,
                     technical=list(NCYCLES=4000)) 
#mg1 <- multipleGroup(d_mat, 1, group = sex_group, 
#                     invariance = c("free_var", "free_means"))

@@

# constrain slopes, allow varying group means and variances -- very slow!
mg4 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance =c('slopes','free_means', 'free_var'), verbose = T)

@@

anova(mg1, mg2) # AIC and BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer mg3, with varying group means
anova(mg3, mg4) # AIC/BIC prefer mg3, with varying group means but equal variance

anova(mod_sc1, mod_fullconstrain) # mod_fullconstrain preferred by BIC, sc1 by AIC
anova(m_sc2, mod_fullconstrain) # mod_fullconstrain preferred by BIC, sc2 by AIC

anova(mod_fullconstrain, mg2) # AIC and BIC both prefer slope-invariant model

@@

coef(mg3)[[2]]$GroupPars # Female intercept = 0.047 (Male = 0)

@@

save(file = here("data/eng_wg_comp_sex_sexDIF.Rds"), 
     mg1, mg2, mg3, mg4, m_sc2, mod_sc1, mod_fullconstrain) 
```

@@

# same item estimates with/without free_means? 
cor(mg2_df$d_f, mg3_df$d_f) # .999
cor(mg2_df$d_m, mg3_df$d_m) # .999

anova(mg1, mg2) # AIC and BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer mg3, with varying group means
```

@@

### Anchor items

For our DIF analysis, let's pick 10 anchor items that show the smallest difficulty difference in the slope-invariant free group means model (mg3).

```{r}
sex_anchors = mg3_df %>% arrange(d_diff_abs) %>% head(10) %>% 
  relocate(a1, .after=definition) %>%
  relocate(d_m, .after=d_f)

sex_anchors %>% kable(digits = 3, format = "html", table.attr = "style='width:40%;'")
```

@@

d_demo <- d_demo %>% left_join(fscores_mg3 %>% 
                                 mutate(data_id = as.numeric(data_id)))

@@

## SES-related DIF

Now we will look for SES-based DIF, using mother's education as a proxy for SES. 
Here's the distribution of mom_ed in our data:

```{r, mom-ed-tab, echo=F}
table(d_demo$mom_ed) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

We (arbitrarily) define high SES as their mother having at least some college (N=`r length(which(d_demo$ses_group=="high"))`), and low SES as having at most secondary education (N=`r length(which(d_demo$ses_group=="low"))`).

@@

mg1 <- multipleGroup(d_mat, 1, group = ses_group, SE = T, technical=list(NCYCLES=4000)) 
mg2 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)
mg3 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = c('slopes', 'free_means'), verbose = T)

anova(mg1, mg2) # AIC/BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer slope-invariant model without varying group mean

@@

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg1, mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain)
```

@@

coef(mg3)[[2]]$GroupPars # low-SES group (high is ref) intercept = .25
```

@@

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg2_df$d_diff_abs) + sd(mg2_df$d_diff_abs)
big_dif = mg2_df[which(mg2_df$d_diff_abs > thresh),]

@@

```{r, fig.width=8, fig.height=6}
ggplot(mg2_df, 
       aes(x = d_lo, y = d_hi)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for low-SES") + 
  ylab("Item Ease for high-SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

### Anchor items

For our DIF analysis, let's pick 10 anchor items that show the smallest difficulty difference in the slope-invariant model (mg2).

```{r ses-anchors, echo=F}
ses_anchors = mg2_df %>% arrange(d_diff_abs) %>% head(10) %>% 
  relocate(a1, .after=definition) %>%
  relocate(d_hi, .after=a1)

ses_anchors %>% kable(digits = 3, format = "html", table.attr = "style='width:40%;'")
```
